### 2025-05-17 Code/scripts_run.Rmd

---
title: "Script Run"
output: html_document
---

# Setup

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source("00_setup.R")
```

# Transport Utilities

```{r transport-utils, echo=FALSE, message=FALSE, warning=FALSE}
source("01_transport_utils.R")
```

# Generate Data

```{r generate-data, echo=FALSE, message=FALSE, warning=FALSE}
source("02_generate_data.R")
```

# Parametric Baseline

```{r param-baseline, echo=FALSE, message=FALSE, warning=FALSE}
source("03_param_baseline.R")
```