### 2025-06-24 analysis_pipeline.Rmd

---
title: "Analysis Pipeline"
output: html_document
---

```{r setup, include=FALSE}
source("00_globals.R")
source("01_data_generation.R")
source("02_split.R")
source("models/true_model.R")
source("models/trtf_model.R")
source("models/ks_model.R")
source("04_evaluation.R")
source("EDA.R")
```

## Data Preparation

```{r}
set.seed(1)
perm <- c(3, 4, 1, 2)
config <- list(
  list(distr = "norm", parm = NULL),
  list(distr = "exp",  parm = function(d) list(rate = softplus(d$X1))),
  list(distr = "beta", parm = function(d) list(shape1 = softplus(d$X2), shape2 = softplus(d$X1))),
  list(distr = "gamma", parm = function(d) list(shape = softplus(d$X3), scale = softplus(d$X2)))
)
prep <- prepare_data(50, config, perm)
```

## Fit Models

```{r}
mods <- fit_models(prep$S, config)
```

## Log-Likelihood Tables

```{r}
tab <- calc_loglik_tables(mods, config)
print(tab)
```

## Scatter Data and Plots

```{r}
scat <- make_scatter_data(mods, prep$S)
scat_plots <- plot_scatter(scat)
```

```{r}
for(p in scat_plots) replayPlot(p)
```

## Optional Full Pipeline

```{r}
# run_pipeline(50, config, perm)
```