### 2025-09-20 report/chapter05.Rnw

% LaTeX file for Chapter 05
<<'preamble05',include=FALSE>>=
library(knitr)
opts_chunk$set(
    fig.path='figure/ch05_fig', 
    self.contained=FALSE,
    cache=TRUE
) 
@

\chapter{References}\label{ch:refs}

\bibliographystyle{mywiley}
\bibliography{biblio}