### 2025-09-20 report/chapterA1.Rnw

% LaTeX file for Chapter 01
<<'preambleA1',include=FALSE>>=
library(knitr)
opts_chunk$set(
    fig.path='figure/cha1_fig', 
    self.contained=FALSE,
    cache=TRUE
)

@

\chapter{Appendix}\label{ch:appendix}

Maybe some R code here, probably a \rr{sessionInfo()}