### 2025-09-22 report/MSc_Report.Rnw
.

\documentclass[11pt,a4paper,twoside]{book}
\input{header.sty}   % packages, layout and standard macros

\begin{document}
\input{TitelblattMABiostatistics}

<<'main-setup',include=FALSE,purl=FALSE>>=
library(knitr)
opts_chunk$set(
    concordance=TRUE,
    child.command = 'include',
    cache=TRUE)
@

\graphicspath{{./figure/}}
\DeclareGraphicsExtensions{.pdf,.png}
\frontmatter
\pagenumbering{roman}
\setcounter{tocdepth}{1}

\chapter*{Abstract}
% Checklist (Abstract): [x] integrate supplied summary [ ] confirm PDF rebuild

This thesis studies multivariate density estimation for tabular data within a transport-based evaluation framework.
All models standardize features and learn a monotone lower-triangular map $S \colon \mathbf{u} \to \mathbf{z}$ to a Gaussian reference distribution.
The estimators accumulate exact log-Jacobians in standardized space with an affine correction for reporting on the data scale.
We compare separable triangular transport maps (TTM-Sep), additive-predictor Transformation Random Forests (TRTF), and copula baselines under the shared preprocessing pipeline.
Additive TRTF induces the same separable triangular likelihood as TTM-Sep via the probability integral transform.
The resulting estimator yields exact likelihoods, transparent conditionals, and linear per-sample cost but restricts context-dependent shape.

Benchmarks include synthetic Half-Moon and a four-dimensional autoregressive generator, plus the UCI POWER, GAS, HEPMASS, and MINIBOONE datasets.
For Half-Moon with $n = 250$, joint negative log-likelihoods (nats) were $1.71$ for TRTF, $1.93$ for TTM-Sep, and $1.54$ for the copula baseline.
For the four-dimensional generator, TRTF achieved $4.53$, TTM-Sep $5.66$, and the copula baseline $5.45$.
Permutation averages confirmed ordering sensitivity for triangular maps.
On MINIBOONE with $K = 43$ and $N = 2{,}500$, TRTF reached $-30.01$ test log-likelihood, whereas flow baselines remained near $-11.7$.

Findings indicate that TRTF narrows the gap to oracle likelihoods on low-dimensional synthetic data.
The method trails modern flows on high-dimensional real data because separability limits context-dependent shape.
Copula baselines remain competitive in low dimensions.
This thesis records guidance and limitations, including ordering effects and fragility in sparse regimes.

\tableofcontents
\setkeys{Gin}{width=.8\textwidth}

\cleardoublepage
\mainmatter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<'child-chapter01', child='chapter01.Rnw'>>=
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<'child-chapter02', child='chapter02.Rnw'>>=
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<'child-chapter03', child='chapter03.Rnw'>>=
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<'child-chapter04', child='chapter04.Rnw'>>=
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<'child-chapter05', child='chapter05.Rnw'>>=
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix
<<'child-chapterA1', child='chapterA1.Rnw'>>=
@

\end{document}