### 2025-09-22 report/chapter04.Rnw

% LaTeX file for Chapter 04
<<'preamble04',include=FALSE>>=
library(knitr)
opts_chunk$set(
    fig.path = 'figure/ch04_fig',
    self.contained = FALSE,
    cache = FALSE
)
@

\chapter{Interpretation and Conclusion}\label{ch:conclusion}

This chapter synthesizes the empirical evidence gathered in Chapter~\ref{ch:dataanalysis}, interprets the behaviour of the estimators within the unified transport frame, and prepares the concluding guidance that follows. We retain the shared preprocessing, likelihood conventions, and diagnostic procedures so that numerical comparisons remain meaningful across synthetic and real datasets.

\section{Interpretation of Results}\label{sec:interpretation-results}
This section interprets the empirical evidence under the unified transport frame. We focus on TRTF-AP, TTM-Sep, and copula baselines evaluated with matched preprocessing, metrics, and units. Synthetic studies report NLL, real datasets report LL, and all values appear in nats with the shared affine correction. These commitments keep objectives, diagnostics, and compute interoperable across estimators.

TRTF-AP often leads within the separable family because additive predictors shift conditional location while the underlying monotone shapes remain stable. The likelihood identities equate TRTF-AP with separable triangular maps, so observed gaps arise from how each estimator realises context shifts and stabilises derivatives. On Half-Moon ($K=2$), TRTF-AP achieved an NLL of $1.71$ while TTM-Sep reached $1.93$, and the first coordinate remained the main source of residual error. Table~\ref{tab:halfmoon-nll} records the per-dimension decomposition and associated uncertainty bands, showing that location adjustments dominate the remaining discrepancies when separability holds approximately in low dimensions.

The four-dimensional generator sharpens this interpretation by isolating coordinates with different conditional structure. TRTF-AP matched the exponential coordinate with an NLL of $1.51$ compared with $1.49$ for the oracle, whereas TTM-Sep over-penalised that coordinate. The beta coordinate produced negative NLLs for the oracles because valid densities can exceed one on $(0,1)$; TRTF-AP approached those values at $-0.25$. The gamma coordinate remained the most challenging, with TRTF-AP at $1.99$ and TTM-Sep at $2.41$. Joint sums favoured TRTF-AP at $4.53$ versus $5.66$, consistent with concentrated gains on location-dominated coordinates. Table~\ref{tab:autoregressive-nll} lists these values, and Figure~\ref{fig:autoregressive-joint-calibration} visualises the residual curvature relative to the identity line.

These comparisons reveal where separability fails to adapt to context-dependent shape. Under a separable map, conditional variance, skewness, and modality remain fixed after the location shift. Probability-integral-transform diagnostics display U-shaped or inverted-U patterns when dispersion misaligns, indicating under- or over-dispersion rather than pure location error. The calibration plots corroborate the per-dimension NLLs and localise remaining structure to the beta and gamma coordinates, where separability is least appropriate. Figure~\ref{fig:autoregressive-joint-calibration} summarises these deviations under the canonical ordering.

Ordering sensitivity stems from finite parameterisations, not from the triangular theory itself. A Knothe--Rosenblatt rearrangement exists for any order, yet limited bases introduce anisotropy that affects fit. Averaging over all $24$ permutations yielded joint NLLs of $4.65$ for TRTF-AP and $5.62$ for TTM-Sep, leaving a $0.97$~nat gap that persisted despite order changes, while the copula baseline averaged $5.45$. Table~\ref{tab:autoregressive-perm} consolidates these permutation-averaged results and underlines the value of data-driven orderings when available.

Small-sample regimes amplified numerical fragility through the log-Jacobian accumulation. TRTF-AP decreased from $38.18$ to $4.64$ joint NLL as $n$ grew from $25$ to $250$, reflecting stabilisation with additional data. TTM-Sep spiked to $6{,}829.45$ at $n=25$ and dropped to $5.61$ at $n=250$, indicating overflow rather than intrinsic misfit. Table~\ref{tab:autoregressive-perm-avg} reports these trajectories, and Section~\ref{sec:models-implementation} documents the derivative clipping and ridge penalties that mitigate this failure mode when samples are scarce.

High dimensionality converts small calibration errors into large likelihood gaps because the triangular determinant accumulates coordinate-wise discrepancies. MINIBOONE with $K=43$ illustrates this accumulation: published flows achieved LL values between $-15.59$ and $-11.68$, whereas TRTF-AP reached $-29.88$ under the shared preprocessing. Table~\ref{tab:uci-loglik} positions TRTF-AP beside the flow baselines and shows that the improvement over the Gaussian reference remains clear even though an approximately $18$~nat gap persists to the strongest flow.

Compute profiles contextualise these accuracy patterns without changing the qualitative ranking at large $K$. At $N=1000$, TRTF-AP required $115$~s on POWER, $138$~s on GAS, $153$~s on HEPMASS, and $202$~s on MINIBOONE, matching the near-linear growth in the training budget and $\mathcal{O}(K)$ evaluation cost. Table~\ref{tab:real-compute} summarises these wall-clock measurements and highlights that separable estimators remain practical in moderate dimensions, yet accuracy dominates the choice once $K \approx 40$.

Taken together, the transport frame delineates when separability suffices and when richer models become necessary. TRTF-AP leads within the separable family when location shifts capture most structure, exhibits ordering sensitivity only through finite bases, and stabilises with modest sample sizes under the safeguards of Section~\ref{sec:models-implementation}. Performance degrades in high dimensions where shape changes and interactions matter, at which point non-separable models offer clear likelihood gains. These conclusions motivate the guidance that will follow in the concluding subsection of this chapter.

\section{Conclusions, Limitations, and Outlook}\label{sec:conclusion-outlook}
We conclude that separable transports remain competitive when conditional location shifts dominate and dimensionality is modest. TRTF-AP led TTM-Sep on Half-Moon ($1.71$ versus $1.93$~NLL) and matched the exponential coordinate in the four-dimensional generator, supporting this interpretation. Conditional decompositions and calibration plots indicate that residual error concentrates in context-dependent shapes, particularly on the beta and gamma components. These findings align with permutation averages that favour TRTF-AP and quantify finite-basis anisotropy. Tables~\ref{tab:halfmoon-nll}--\ref{tab:autoregressive-perm} together with Figure~\ref{fig:autoregressive-joint-calibration} document this evidence under the shared protocol.

Performance on MINIBOONE reveals the cost of separability at higher dimension. TRTF-AP improved the Gaussian reference yet remained about $18$~nats behind the best published flow, consistent with accumulated Jacobian error across $43$ coordinates. POWER exhibited the opposite regime, where TRTF-AP surpassed the reported flow baselines under identical preprocessing. These contrasts suggest that conditional shape and dimension jointly determine whether separable structure suffices. Table~\ref{tab:uci-loglik} reports these comparisons in a common unit.

Compute profiles remained practical and scaled near-linearly with the training budget. Training plus evaluation required $115$~s at $N=1000$ on POWER and $202$~s on MINIBOONE, with longer totals at $N=2500$ that preserved the same trend. These measurements keep separable transports viable for exploratory analysis and model diagnostics. Table~\ref{tab:real-compute} records the budgeted timings and the shared pipeline settings.

Several limitations qualify these conclusions. Separable maps fix conditional shape and therefore cannot resolve heteroskedasticity or conditional multimodality. Ordering remained a material source of variance under finite bases, as shown by the $0.97$~nat permutation gap despite stable rankings at moderate sample sizes. Small-sample regimes created numerical fragility through steep log-Jacobian terms, which clipping and ridge regularisation mitigate but do not eliminate. Real-data tables still contain missing GAS and HEPMASS entries, and single-seed settings persist for some runs, limiting external comparability. Tables~\ref{tab:autoregressive-perm}--\ref{tab:real-compute} catalogue these caveats within the standardised protocol.

The outlook follows directly from the evidence. Data-driven orderings are likely to reduce anisotropy without abandoning the lower-triangular map. Low-rank cross-terms in triangular transports and non-additive predictors in TRTF may adapt conditional shapes while preserving monotonicity, exact inversion, and linear per-sample evaluation. Expanded calibration reporting, including probability-integral-transform summaries and Kolmogorov--Smirnov distances, should remain part of any deployment-grade evaluation. Completing GAS and HEPMASS under the same protocol will improve generality and sharpen the accuracy-versus-compute trade-off. These steps target smaller likelihood gaps on high-$K$ datasets while retaining the interpretability and reproducibility provided by the transport frame.